---
title: "Introduction to ConvNets with Keras and R"
output: html_notebook
---

Load packages:

```{r, message=FALSE}
library(keras)
library(tidyverse)
library(jsonlite)
library(abind)
library(gridExtra)
library(pracma)
source('utils.R')
```

We will use a dataset of 2800 satellite pictures. Every row contains information about one photo (80-pixel height, 80-pixel width, 3 colors - RGB color space). To input data into a Keras model, we need to transform it into a 4-dimensional array (index of sample, height, width, colors). Every picture is associated with a label that could be equal 1 for a ship and 0 for non-ship object. Read the data:

```{r}
ships_json <- fromJSON("shipsnet.json")[1:2]
names(ships_json)
dim(ships_json$data)
length(ships_json$labels)
```

TASK1: Transform the image into 4 dimensional array:

```{r}
sample_image <- ships_json$data[1,]
```

From `sample_image` select pixel values corresponding to RED, GREEN and BLUE channels and reshape them into matrices (dims = height X width)

```{r}
r <- matrix(sample_image[1:?], ?, ?, byrow = TRUE) / 255
g <- matrix(sample_image[?:?], ?, ?, byrow = TRUE) / 255
b <- matrix(sample_image[?:?], ?, ?, byrow = TRUE) / 255
```

Bind channels into 3-dimensional array:

```{r}
sample_image <- array(c(?, ?, ?), dim = c(?, ?, ?))
```

Plot `sample_image`:

```{r}
plot_sample_image(sample_image, show_layers = TRUE)
```

Sometimes it is usefull to do some geometric transformation (rotation, translation, thickening, blurring etc.) to enlarge training set on an image to create new one. For example, in `R` we can use `rot90` function from the `pracma` package to create images rotated by 90, 180, or 270 degrees. Rotate the images:

```{r}
sample_image_rot90 <- array(c(rot90(r, 1), ?, ?), dim = c(80, 80, 3))
sample_image_rot180 <- array(?, ?, ?), dim = c(80, 80, 3))
sample_image_rot270 <- array(?, ?, ?), dim = c(80, 80, 3))

grid.arrange(plot_sample_image(sample_image),
             plot_sample_image(sample_image_rot90),
             plot_sample_image(sample_image_rot180),
             plot_sample_image(sample_image_rot270),
             ncol = 2, nrow = 2)
```

Finally, apply it to all images in the dataset:

```{r}
ships_data <- ships_json$data %>% apply(., 1, function(x) {
  r <- matrix(x[1:6400], 80, 80, byrow = TRUE) / 255
  g <- matrix(x[6401:12800], 80, 80, byrow = TRUE) / 255
  b <- matrix(x[12801:19200], 80, 80, byrow = TRUE) / 255
  list(array(c(r,g,b), dim = c(80, 80, 3)),
       array(c(rot90(r, 1), rot90(g, 1), rot90(b, 1)), dim = c(80, 80, 3)),
       array(c(rot90(r, 2), rot90(g, 2), rot90(b, 2)), dim = c(80, 80, 3)),
       array(c(rot90(r, 3), rot90(g, 3), rot90(b, 3)), dim = c(80, 80, 3)))
}) %>% do.call(c, .) %>% abind(., along = 4) %>% aperm(c(4, 1, 2, 3))
ships_labels <- ships_json$labels %>% map(~ rep(.x, 4)) %>%
  unlist() %>% to_categorical(2)
dim(ships_data)
dim(ships_labels)
```

Plot sample images:

```{r}
plot_sample_images(ships_data, ships_labels, seed = 226)
```

Split the data into training and test set:

```{r}
set.seed(1234)
indexes <- sample(seq(1, 2800 * 4, 4), 0.7 * 2800) %>% map(~ .x + 0:3) %>% unlist()
train <- list(data = ships_data[indexes,,,], labels = ships_labels[indexes,])
test <- list(data = ships_data[-indexes,,,], labels = ships_labels[-indexes,])
rm(ships_json, ships_data)
```

TASK2: Build CNN for image classification:

Initialize `model1`:

```{r}
model1 <- keras_model_sequential()
summary(model1)
```

Add first convolution layer to `model1`. Specify the input shape, nr of filters (32), kernel_size (3x3), strides (1x1) and activation function (relu). Note that objects in Keras are `modified in-place` so there’s no need for consecutive assignment:

```{r}
model1 %>%
  layer_conv_2d(
    input_shape = c(?, ?, ?),
    filter = ?, kernel_size = c(?, ?), strides = c(?, ?),
    activation = ?)
summary(model1)
```

Add another layers to `model1`:
1. Max pooling layer (layer_max_pooling_2d) with pool size 2x2 and strides 2x2,
2. Convolution layer (layer_conv_2d) with 64 filters, kernel 3x3, strides 1x1, relu activation,
3. Max pooling layer (layer_max_pooling_2d) with pool size 2x2 and strides 2x2,
4. Flatten the resulting activation map (layer_flatten),
5. Output layer giving the probabilities of a ship in an image (layer_dense). Use softmax activation.

```{r}
model1 %>%
  ?
summary(model1)
```

Configure `model1` for training:

```{r}
model1 %>% compile(
  loss = "categorical_crossentropy",
  optimizer = optimizer_sgd(lr = 0.0001, decay = 1e-6),
  metrics = "accuracy"
)
```

Fit `model1`:

```{r error=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
ships_fit1 <- model1 %>% fit(train[[1]], train[[2]], epochs = 20, batch_size = 32,
                             validation_split = 0.2,
                             callbacks = callback_tensorboard("logs/model1"))
```

Predict on test set:

```{r}
predicted_probs <- model1 %>%
  predict_proba(test[[1]])

head(predicted_probs)

model1 %>% evaluate(test[[1]], test[[2]])
```

Save/load the model:

```{r}
save_model_hdf5(model1, "model1.hdf5")
new_model <- load_model_hdf5("model1.hdf5")
```

TASK3: Build, train and fit your own model:
I. Architecture:
1. Start with 2d convolution (32 filters, 3x3 kernel, relu activation)
2. Add another 2d convolution
3. Add max pooling layer (2x2 pool size)
4. Add dropout layer (rate 0.25)
5. Add another 2d convolution (64 filters)
6. Add another 2d convolution (64 filters)
7. Add max pooling layer (2x2 pool size)
8. Add dropout layer (rate 0.25)
9. Add dense layer (512 units)
II. Configure model for training:
1. Instead of SGD, use ADAMAX optimizer
III. Fit the model
1. Use TensorBoard callback ( dir = "logs/model2")

```{r, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
model2 <- ?
model2 %>%
  ?

model2 %>% compile(
  loss = "categorical_crossentropy",
  optimizer = ?,
  metrics = "accuracy"
)

ships_fit2 <- model2 %>% ?

predicted_probs <- model2 %>%
  predict_proba(test[[1]])

head(predicted_probs)

model2 %>% evaluate(test[[1]], test[[2]])
```

Compare models:

```{r, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
tensorboard("logs")
```

